---
title: "Multivariate Adaptive Regression Splines (MARS)"
output:
  pdf_document: default
  word_document: default
  html_document:
    df_print: paged
vignette: "%\\VignetteIndexEntry{Mars} %\\VignetteEngine{knitr::rmarkdown} %\\VignetteEncoding{UTF-8}\n"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Github Page
https://github.com/DustinPoon/STAT360Project

# Description

Using methods from Friedman's paper "Multivariate Adaptive Regression Splines" to construct a regression model.

# Usage
```{r eval=FALSE}
mars(formula, data, control = NULL)
```

# Arguments

  *formula*   an R formula.
  
  *data*      a data frame containing the data for the model.
  
  *control*   an object of class 'mars.control'.

# Details

Jerome H. Friedman introduced a type of regression analysis called Multivariate Adaptive Regression Spline (MARS). MARS is a non-parametric regression technique that automatically models nonlinearities and interactions between variables. MARS begins by recursively partitioning the data into two subsets and fitting a linear regression model to the regions. The predictor variable is where the subsetting occurs and is chosen by the forward stepwise method, which selects the best predictor variable. Again, repeated for each subset. A new predictor variable is added to the model, using the residuals from the two models. The process is repeated until the Mmax of basis functions is reached. The resulting model consists of the selected basis functions and their formulas, the names of the predictor variables, and the fitted linear model object.

\newpage

## Key Aspects in Mars Algorithm

### Forward Step-wise
The forward step-wise algorithm recursively partitions the predictor variable space into sub-regions by adding
pairs of basis functions, each representing a sub-region. The algorithm selects the optimal basis function to split
on by iterating over four loops, with each iteration selecting a variable and split point based on the GCV criterion.
The basis functions are recorded in a list called Bfuncs and a matrix called B.

### Backward Step-wise
The backward step-wise algorithm removes terms one by one, selecting the least effective term at each step until
it finds the best sub-model. Model subsets are compared using the GCV criterion, implemented using two for
loops. The outer loop iterates over all possible model sizes, while the inner loop iterates over all model terms in
the current model and tries removing one basis function at a time. The final model is selected based on the
lowest observed lack of fit value using the GCV criterion.

### Hinge Function
MARS models use hinge functions, which are constructed using a variable x and a split-point (knot) “a.” The
hinge function takes the form max(0, x - a) or max(0, a - x) and can be used to partition data into disjoint regions.
These regions can be treated independently and the hinge functions can be multiplied together to form non-linear
functions.

### Generalized Cross Validation (GCV) Criterion
When selecting models in the MARS algorithm, the Generalized Cross Validation (GCV) criterion is used to
measure “lack of fit” (LOF) instead of the Residual Sum of Squares (RSS). The RSS tends to favor larger models
because it decreases as we add predictors. Cross-Validation is a better approach but can be time-consuming,
which is why we use the GCV criterion as an approximation of Cross-Validation.
The GCV criterion can be calculated using the following formula:

  RSS×(N/(N - C'(M))^2)
N is the number of rows in the dataset.
M is one less than the number of coefficients in the fitted model.
C’(M) is the sum of the hat-values from the fitted model and the smoothing parameter d times M, where d is typically set to 3, according to
Friedman.

# Value

An S3 model of class 'mars'.

# Author(s)

Group DJT: Dustin Poon, James Lee, Tyler Oh

# References

Jerome H. Friedman, The Annals of Statistics, Mar., 1991, Vol. 19, No. 1 (Mar., 1991), pp. 1-67.

# See Also

  *mars.control*    for constructing control object.
  
  *anova.mars*      for the ANOVA table.
  
  *plot.mars*       for plotting the fitted basis function.
  
  *predict.mars*    for Predicting from the fitted model.
  
  *print.mars*      for printing out a mars object.
  
  *summary.mars*    for more detailed summaries of mars object.

*For more detailed explanations, please type ?function_name in your R or Rstudio console. (i.e., ?mars)*

# Examples

## Data

### Wage: R built-in dataset
### iris: The dataset from library(ISLR)
### CarPrice: This dataset is extracted from https://www.kaggle.com/datasets/hellbuoy/car-price-prediction{.underline}

### Setup

```{r setup}
library(Mars)
CarPrice <- read.csv("/Users/tyler/Mars/vignettes/CarPrice_Assignment.csv")
```

### Control

```{r}
mc <- mars.control(Mmax=10)
```

### Wage data

```{r, eval=FALSE}
fit.Wage <- mars(wage ~ age + education, ISLR::Wage, mc)
print.mars(fit.Wage)
summary.mars(fit.Wage)
predict.mars(fit.Wage, newdata = data.frame(age=ISLR::Wage$age,
                                            education = ISLR::Wage$education))
anova.mars(fit.Wage)
plot.mars(fit.Wage)
```

\newpage
### iris Dataset

```{r, eval=FALSE}
fit.iris <- mars(Sepal.Length ~., iris, mc)
print.mars(fit.iris)
summary.mars(fit.iris)
predict.mars(fit.iris, newdata = data.frame(sepal.width = iris$Sepal.Width,
                                            petal.length = iris$Petal.Length,
                                            petal.width = iris$Petal.Width,
                                            Species = iris$Species))
anova.mars(fit.iris)
plot.mars(fit.iris)
```

### CarPrice dataset

```{r, eval=FALSE}
fit.CarPrice <- mars(price ~ citympg + highwaympg + horsepower + fueltype, CarPrice, mc)
print.mars(fit.CarPrice)
summary.mars(fit.CarPrice)
predict.mars(fit.CarPrice, newdata = data.frame(citympg = CarPrice$citympg,
                                                highwaympg = CarPrice$highwaympg,
                                                horsepower = CarPrice$highwaympg,
                                                fueltype = CarPrice$highwaympg))
anova.mars(fit.CarPrice)
plot.mars(fit.CarPrice)
```
